Figures for Paper

### 2020-12-11 paper_figures.Rmd
some model scripts which I will probably delete later

```{r setup_function, echo=FALSE}
####helper funciton that gets the overall model p-value####
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
  # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
    p_value = lmp(model)
    r_squared = round(summary(model)$adj.r.squared,3)
    pretty_plot = broom::tidy(model) %>% 
        mutate(p_star = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  p.value > 0.1 ~ "NS")) %>% 
        filter(p.value < p_value_cutoff) %>%
        mutate(term = fct_reorder(term, -abs(estimate))) %>%
        ggplot(aes(x = term, y = estimate,fill = estimate > 0,show.legend = F)) +
        geom_col() +
        coord_flip() +
        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}")) +
        geom_text(aes(x=term, y= -0.01, label=p_star)) 
}
```
# iPSC splicing

@@

```{r}
junc_colors = c("#648FFF","#fe6101")
ipsc_splicing = fread(file.path(here(),"data","ipsc_splicing_results.csv")

ipsc_splicing %>% 
  mutate(junction_name = glue::glue("{gene_name} - {paste_into_igv_junction}")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ "")) %>% 
  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
                  min.segment.length = unit(0, 'lines')) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  guides(alpha = FALSE, size = FALSE) + 
  scale_fill_manual(values = junc_colors) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab("-Log 10 Test Statistic") + 
  theme(text = element_text(size = 18))
```

@@

rel_rna_cryptic_amount = fread(file.path(here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

 rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE) 
```

# Paient summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi_leaf), n = 2))) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                      stmn2_psi_groups == 2 ~ "High STMN2",
                                      TRUE ~ "No STMN2"))

@@

@@

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals")

@@

####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "FTLD \n non-TDP",
                              "FTLD-TDP")) %>% 
    filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
  dplyr::select(disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
  unique() %>% 
    ggplot(aes(x = disease_group2, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F) + 
    geom_jitter(height = 0,show.legend = F) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 18)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4])

@@

Splitting by the type of junction in ALS
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "ALS \n non-TDP",
                              "ALS-TDP")) %>% 
      filter(tissue_clean %in% c("Cerebellum","Motor Cortex","Temporal Cortex") | grepl("Cord",tissue_clean)) %>% 
  dplyr::select(disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
  unique() %>% 
  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
    ggplot(aes(x = disease_group2, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F) + 
    geom_jitter(height = 0,show.legend = F) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 18)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
  theme(axis.text.x =  element_text(size = 14))
```

@@

@@

## Scatter plot showing the relationship in non-logspace between STMN2 and UNC13A cryptic PSI
```{r}
####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi_leaf",
                      y = "unc13a_cryptic_leaf_psi",
                      add = "reg.line",
                      cor.coef = T) +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ")

####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    mutate(call = fct_relevel(call,"C/C", "C/G", "G/G")) %>%
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi_leaf",
                      y = "unc13a_cryptic_leaf_psi",
                      add = "reg.line",
                      color = "call") +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
    stat_cor(aes(color = call),show.legend = FALSE) +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a"))
```

You can see a clear relationship, which appears to be modulated by the genotype
of rs12973192. This lead us to see if genotype was actually predictive of the 
amount of UNC13A cryptic psi. Because PSI's tend to be log linear, we chose
to model UNC13A in the log10 + 1 (log10(1) is zero) and we selected for tissues
which show TDP-43 pathology, as we assume this is an event which only begins to
show clear after TDP-43 pathology arises.

## Scatter plot showing the relationship in log + 1 between STMN2 and UNC13A cryptic PSI

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1)) %>% 
    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn_2_cryptic_psi_leaf_plusone",
                      y = "log10_unc13a_cryptic_leaf_psi_plusone",
                      color = 'call',fill = 'call',                      
                      add = "reg.line") +
    ylab("Log10 UNC13A Cryptic PSI + 1") + 
    xlab("Log10 STMN2 Cryptic PSI + 1 ") +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) + 
    scale_fill_manual(values = c("#88CCEE","#44AA99","#105e2a"))

@@

```{r}
####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
    filter(tissue_clean%in% c("Cervical_Spinal_Cord",
                              "Motor_Cortex",
                              "Frontal_Cortex","Temporal_Cortex")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(cryptic_psi > 0 ) %>% 
    filter(stmn_2_cryptic_psi > 0 ) %>% 
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi",
                      y = "unc13a_cryptic_leaf_psi",
                      color = 'call',                      
                      add = "reg.line") +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
    facet_wrap(~tissue_clean)

@@

## Modeling UNC13A cryptic using glms and mixed models

This brought us to the idea that we could even potentially model
UNC13A cryptic pathology using genotype and the amount of TDP-43 pathology. 
We've coded the number of G alleles here as a number and are looking into tissue
with TDP-43 pathology(measure with STMN2 cryptic PSI).

new_regress = clean_data_table %>%
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%  
    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1))

another_model2 = lm(log10(unc13a_cryptic_leaf_psi + 1) ~ 
                   log10(stmn_2_cryptic_psi_leaf + 1) * number_g_alleles + pathology + tissue_clean,
              new_regress)

the_pretty_plot = pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI") + ggpubr::theme_pubr()
print(the_pretty_plot)
```

We find that at a cutoff of 0.01 for significant effects, only the interaction
between the number of G alleles and the amount of STMN2 pathology is predictive
of UNC13A cryptic inclusion. the absolute amount of STMN2 pathology does not
vary between genotypes. 
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

ggboxplot(new_regress, x = "call", y = "log10_stmn_2_cryptic_psi_leaf",
          color = "call", palette = "jco")+
  stat_compare_means() +
  stat_compare_means(comparisons = genotype_comparisons)

ggboxplot(new_regress, x = "call", y = "stmn_2_cryptic_psi_leaf",
          color = "call", palette = "jco")+
  stat_compare_means() +
  stat_compare_means(comparisons = genotype_comparisons)

new_regress %>% 
    ggplot() + 
    stat_ecdf(aes(color = call, x = log10_stmn_2_cryptic_psi_leaf))
```

We see that detection rate by genotype varies.

>>>>

# iPSC splicing

@@

```{r}
junc_colors = c("#648FFF","#fe6101")
ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))

ipsc_splicing %>% 
  mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ gene_name,
                                    T ~ "")) %>% 
  mutate(junction_name = make.unique(junction_name,sep = ".")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ "")) %>% 
  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
                  min.segment.length = unit(0, 'lines')) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  guides(alpha = FALSE, size = FALSE) + 
  scale_fill_manual(values = junc_colors) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab("-Log 10 Test Statistic") + 
  theme(text = element_text(size = 18))
```

@@

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

unca = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)

stmn2 = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "stmn_2_cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("STMN2 Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B Intron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE)

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Intron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE)

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
ggarrange(unca,stmn2)
```

# Paient summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi_leaf), n = 2))) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                      stmn2_psi_groups == 2 ~ "High STMN2",
                                      TRUE ~ "No STMN2"))

@@

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```

@@

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals")
```

# Detection UNC13A by tissue

```{r}
boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n {tissue_clean} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = disease_group2, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  facet_wrap(~tissue_clean) +
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  
```

# Detection UNC13A by tissue and Genotype

```{r}
boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
  left_join(clean_data_table %>% dplyr::select(participant_id,call)) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean,call) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ggplot() +
  geom_col(aes(x = disease_group2, y = detection_rate,fill = call),position = "dodge") + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  facet_wrap(~tissue_clean) +
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  
```

@@

####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "FTLD \n non-TDP",
                              "FTLD-TDP")) %>% 
    filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
  dplyr::select(disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
  unique() %>% 
    ggplot(aes(x = disease_group2, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F) + 
    geom_jitter(height = 0,show.legend = F) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A cryptic \n reads per million") +
    xlab("") +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
      theme(text = element_text(size = 18))

@@

Splitting by the type of junction in ALS
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "ALS \n non-TDP",
                              "ALS-TDP")) %>% 
      filter(tissue_clean %in% c("Cerebellum","Motor Cortex","Temporal Cortex") | grepl("Cord",tissue_clean)) %>% 
  dplyr::select(disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
  unique() %>% 
  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
    ggplot(aes(x = disease_group2, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F) + 
    geom_jitter(height = 0,show.legend = F) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A cryptic \n reads per million") +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  
```

@@

# Correlation between STMN2 Cryptic PSI and UNC13A TPM

Separated by tissue only in ALS/FTLD-TDP
```{r}

clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = "free_y")
```

No Tissue separation

```{r}
clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  ylim(0,50)
```

```{r}
#tissue separated
clean_data_table %>% 
      filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A Cryptic PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free')

clean_data_table %>% 
      filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(label.y = 38,size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("STMN2 Cryptic PSI") + 
  ylim(0,40) +
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18))

@@

# Relationship between STMN2 and UNC13A cryptic PSI in patients
```{r}
####scatter plot showing the correlation in non-log space in STMN2 and cryptic PSI####
clean_data_table %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A Cryptic PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,call,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  data.table::melt() %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  mutate(variable = ifelse(variable == "stmn_2_cryptic_psi_leaf", "STMN2", "UNC13A")) %>% 
  ggbarplot( x = "tissue_clean", 
               y = "value", 
               color = 'variable',
              position = position_dodge(0.8),
              add = c("mean_se","jitter")) +
  ylab("Cryptic PSI event") + 
  facet_wrap(~variable,scales = "free") + 
  scale_color_manual(values = colorblind_pal()(4)[1:3]) +
  theme(legend.position = 'none') + 
  xlab("") +
  theme(text = element_text(size = 18))

@@

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
    filter(call != "C/G") %>% 
    ggpubr::ggscatter(., 
                      x = "stmn_2_cryptic_psi",
                      y = "unc13a_cryptic_leaf_psi",
                      color = 'call',                      
                      add = "reg.line",
                      ) +
    ylab("UNC13A Cryptic PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes(                      x = stmn_2_cryptic_psi,
                      y = unc13a_cryptic_leaf_psi,
                      color = call ),show.legend = FALSE) + 
  scale_color_manual(values = c("#88CCEE","#105e2a")) +
  theme(text = element_text(size = 18))

@@

# Detection rate by genotype

## Barplot of detection by genotype
